---
title: "Week 2 Assignment"
author: "Beata Vekeriotaite"
date: "3/7/2022"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(ensembldb)
  library(GenomicRanges)
  library(epiwraps)
})
```

# Find and download the mouse (Mus Musculus) EnsDb object, version 102, genome build GRCm38

```{r}

ah <- AnnotationHub()

# to query the AH programmatically:

ensdb_mouse <- query (ah, c("Mus Musculus", "EnsDb", "GRCm38", "102"))

# to query the AH programmatically:

# to browse the AnnotationHub interactively:

# display(ah)

ensdb <- ah[["AH89211"]]

```

# Find and download the mouse (Mus Musculus) genome sequence (dna_sm) in TwoBit/2bit format for GRCm38

```{r}

dna_sm_grcm38 <- ah[["AH88477"]]

```


# Find and download the drosophila melanogaster genome sequence ( dna_sm ) in TwoBit/2bit format for BDGP6

```{r}

dna_sm_bdgp6 <- ah[["AH99424"]]

```


# Using the mouse EnsDb, find the following:

## How many different ensembl gene IDs and gene symbols are there for protein-coding genes?

```{r}

# protein-coding genes

gene_IDs <- transcripts (ensdb, filter = GeneBiotypeFilter("protein_coding"))

length(unique(gene_IDs$gene_id)) + length(unique(gene_IDs$gene_name))

# protein-coding transcripts

txs <- transcripts(ensdb, columns=c("tx_id", "tx_biotype", "gene_id", "gene_name"), filter =TxBiotypeFilter("protein_coding"))

length(unique(txs$gene_id)) + length(unique(txs$gene_name))


## not sure which one I was supposed to go with here. 

```

## Plot the distribution (histogram) of how many exons protein-coding genes have

```{r}
txs <- transcripts(ensdb)

exs_per_tx <- exonsBy(ensdb, column=c("tx_id","tx_biotype"), 
                    filter=TxBiotypeFilter("protein_coding"))

total_exons_per_transcript <- lengths(exs_per_tx)

hist(total_exons_per_transcript, xlim = c(0,50), breaks = 300, ylim = c(0,9500))



```


## Plot the distribution of the (spliced) length of protein-coding transcripts

```{r}

protein_coding_txs <- transcripts (ensdb, filter = TxBiotypeFilter("protein_coding"))

plot(width(protein_coding_txs))

```



